---
title: "General R Commands"
output: 
  html_document: 
    theme: cerulean
    highlight: tango
    css: styles.css
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

----

Add custom notes to the [My Quick R Notes](#quickR) section below.

----

### data( ) {#data}

<div style="padding-left:15px;">
**Console** Help Command: `?data`

`data()`

* This command lists the available datasets in R. Information about each dataset can then be viewed using the `?nameofthedataset` command.

**Example Code**

```{r, eval=FALSE, comment=NA}
#Run these codes in your Console to see what they do.

data()
?cars #Brings up the help file for the cars dataset.
?CO2 #Brings up the help file for the CO2 dataset.
```

</div>
<br />
<hr />

### head( ) {#head}

<div style="padding-left:15px;">
**Console** Help Command: `?head`

`head(object)`

There is a similar function called `tail(object)`

* `object` is typically the name of a `data.frame` object in R, i.e., a dataset.

**Example Code**

```{r, eval=FALSE, comment='>'}
#Run these codes in your Console to see what they do.

head(cars) #Shows the first 6 rows by default.
head(cars,11) #Shows the first 11 rows.
tail(CO2) #Shows the last 6 rows by default.
tail(CO2, 10) #Shows the last 10 rows.
```

</div>
<br />
<hr />


### View( ) {#view}

<div style="padding-left:15px;">
**Console** Help Command: `?View`

`View(object)`

* `object` is typically the name of a `data.frame` object in R, i.e., a dataset.

**Example Code**

```{r, eval=FALSE, comment='>'}
#Run these codes in your Console to see what they do.

View(cars) #Shows the entire cars dataset
View(CO2) #Shows the entire CO2 dataset
```

</div>
<br />
<hr />



### $ {#dollar}

<div style="padding-left:15px;">
`DataSetName$ColumnName`

* The `$` operator allows you to access the individual columns of a dataset.

**Example Code**

```{r, eval=FALSE, comment=NA}
#Run these codes in your Console to see what they do.

View(cars) #Shows the entire cars dataset
cars$speed #Access the speed column from the cars dataset
cars$dist #Access the dist column from the cars dataset
```

</div>
<br />
<hr />


### with( ) {#with}

<div style="padding-left:15px;">
**Console** Help Command: `?with`

`with(datasetName, commands)`

* This function allows you to not have to use `$` to access the columns of a dataset. It is useful sometimes, but not always.
* `datasetName` is the name of a dataset like `cars` or `CO2`.
* `commands` are any R Commands that need to be performed using the variables within the specified dataset.

**Example Code**

```{r, eval=FALSE, comment=NA}
#Run these codes in your Console to see what they do.

with(cars, mean(dist)) #same thing as: mean(cars$dist)
with(cars, cor(dist, speed)) #same as: cor(cars$dist, cars$speed)
```

</div>
<br />
<hr />


### subset( ) {#subset}

<div style="padding-left:15px;">
**Console** Help Command: `?subset`

`subset(NameOfDataset, condition)`

* `NameOfDataset` is the name of a dataset, like `cars` or `CO2`.
* `condition` is some logical expression that uses any of the following logical connectors.

<div style="padding-left:30px;">
| **Logical Expression** | **Syntax** | 
|--------------------|--------|
| Equals | `==` |
| Not Equal | `!=` |
| Less Than | `<` |
| Less Then or Equal to | `<=` |
| Greater Than | `>` |
| Greater Than or Equal to | `>=` |
| AND | `&` | 
| OR | `|` |
| IN | `%in%` |
| NOT | `!` |
</div>

**Example Code**

```{r, eval=FALSE, comment=NA}
#Run these codes in your Console to see what they do.

subset(CO2, Type=="Quebec")
subset(CO2, Type!="Quebec")
subset(CO2, conc < 175)
subset(CO2, conc <= 175)
subset(CO2, conc > 675)
subset(CO2, conc >= 675)
subset(CO2, Type=="Quebec" & conc <= 175)
subset(CO2, Plant=="Qn1" | Plant=="Mc2")
subset(CO2, Plant %in% c("Qn1","Qc3","Mn2","Mc3"))
subset(CO2, Type !="Quebec")
```

</div>
<br />
<hr />


### <- {#assignment}

<div style="padding-left:15px;">
**Console** Help Command: `none`

`NameYouCreate <- some R commands`

* `<-` (Less than symbol `<` with a hyphen `-`) is called the assignment operator and lets you store the results of the `some R commands` into an object called `NameYouCreate`.
* `NameYouCreate` is any name that begins with a letter, but can use numbers, periods, and underscores thereafter. The name should not contain spaces. 

**Example Code**

```{r, eval=FALSE, comment=NA}
#Run these codes in your Console to see what they do.

aveSpeed <- mean(cars$speed) #Saves the mean into the aveSpeed object
aveSpeed #prints the saved mean to the screen

fastCars <- subset(cars, speed > 20) #Saves the subset into fastCars
View(cars) #Shows the entire cars dataset
View(fastCars) #Shows the entire fastCars dataset

fastCars$NewColumnFTperSEC <- fastCars$speed * 5280 / 3600 #Add new column
View(fastCars) #See the newly expanded fastCars dataset
```

</div>
<br />
<hr />

### c( ) {#cvector}

<div style="padding-left:15px;">
**Console** Help Command: `?c`

`c( )`

* The `c( )` function combines values into a vector or list.

**Example Code**

```{r, eval=FALSE, comment=NA}
#Run these codes in your Console to see what they do.
c(1,3,5)
c(1:5)
c("dog","cat","bird")

animals <- c("dog","cat","bird")
animals

counts <- c(5, 2, 18)
counts
```

</div>
<br />
<hr />


### [ ] {#brackets}

<div style="padding-left:15px;">
**Console** Help Command: `none`

`objectName[ elementNumber ]`

`objectName[ Rows, Columns]` 

* If `objectName` is an object that contains a list of numbers then `[ elementNumber ]` will pull out the specified elements.
* If `objectName` is an object with rows and columns (like a dataset) then the specified `[ Rows , Columns]` can be accessed.

**Example Code**

```{r, eval=FALSE, comment=NA}
#Run these codes in your Console to see what they do.

counts <- c(5, 2, 18)
counts
counts[3]   #third element
counts[2:3]  #second through third elements
counts[c(1,3)] #first and third elements

View(cars)
cars[2:3, ] #second and third rows, all columns
cars[ ,2]   #all rows, second column
cars[2,2]   #second row, second column (single item)
cars[2:3,2] #second and third rows, second column
```

</div>
<br />
<hr />

<!--

### ... {#}

<div style="padding-left:15px;">
**Console** Help Command: `?`

`()`

* 

**Example Code**

```{r, eval=FALSE, comment=NA}
...
```

</div>
<br />
<hr />




<br />
<br />
<br />


-->

<hr />


## My Quick R Notes {#quickR}

Add any R commands you use and want to remember later into the following chunk.

```{r, eval=FALSE}
round(1.55324, 2) #Rounds numbers to 2 decimal places
class(cars) #Shows the R "class" of cars
class(cars$speed) #Shows the R "Class" of cars$speed
str(cars) #Shows the R "structure" of an object


```



<footer></footer>