---
title: "Day Care Fines"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true 
    toc_float: true
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r, include=FALSE}

library(mosaic)
library(DT)
library(pander)
library(plyr)

Late <- read.csv("../Data/Late.csv", header=TRUE)
Late$Period <- mapvalues(Late$Week, 1:20,c(rep("Before", 4), rep("During", 12), rep("After", 4)))
#Late$Status <- cut(Late$Week, c(0,4,16,20), c("s1", "s2", "s3"))
Fine <- subset(Late, Treatment %in% c("Fine"))

Control <- subset(Late, Treatment %in% c("Control"))

```


## Background

Will implementing a fine change how people view their obligations?  A number of parents have been consistently late in picking up their children from day-care centers in Israel.  A study was performed to see what would happen if a fine was introduced.  Details on the study are shown in the links below. 

<a href="javascript:showhide('abstract')">Details of the Day Care Study<span style="font-size:8pt;">&nbsp;(click to view)</span></a>

<div id="abstract" style="display:none;">

<div style="color:#a8a8a8;">
This background is quoted directly from the article ["A Fine is a Price"](http://rady.ucsd.edu/faculty/directory/gneezy/pub/docs/fine.pdf). 
</div>

<div style="padding-left:30px; padding-right:30px;">

There are two types of day-care centers in Israel: private and public. A study was conducted in 10 private day-care centers in the city of Haifa from January to June 1998. All of these centers are located in the same part of town, and there is no important difference among them. During the day children are organized into groups according to age, from 1 to 4 years old. Each day-care center is allowed to hold a maximum of 35 children. In some exceptional cases a few additional children are allowed. The fee for each child is NIS 1,400 per month. (The NIS is the New Israeli Shekel.) At the time of the study, a U.S. dollar was worth approximately NIS 3.68, so the fee was about $380 at that time. 

The contract signed at the beginning of the year states that the day-care
center operates between 0730 and 1600. There is no mention of what happens if parents come late to pick up their children. In particular, before the beginning of the study, there was no fine for coming late. When parents did not come on time, one of the teachers had to wait with the children concerned. Teachers would rotate in this task, which is considered part of the job of a teacher, a fact that is clearly explained when a teacher is hired. Parents rarely came after 1630. 

A natural option [to fix the problem of parents showing up late] is to introduce a fine: every time a parent comes late, [they] will have to pay a fine. Will that reduce the number of parents who come late? If the fine is removed, will things revert back to the way they were originally?

The overall period of the study was 20 weeks. In the first 4 weeks we
simply recorded the number of parents who arrived late each week. At the
beginning of the fifth week, we introduced a fine in six of the 10 day-care centers, which had been selected randomly. The announcement of the fine was made with a note posted on the bulletin board of the day-care center. Parents tend to look at this board every day, since important announcements are posted there. The announcement specified that the fine would be NIS 10 for a delay of 10 minutes or more. The fine was per child; thus, if parents had two children in the center and they came late, they had to pay NIS 20. Payment was made to the principal of the day-care center at the end of the month. Since monthly payments are made to the owner during the year, the fines were added to those amounts. The money was paid to the owner, rather then to the teacher who was staying late (and did not get any additional money). The teachers were informed of the fine but not of the study. Registering the names of parents who came late was a common practice in any case.

At the beginning of the seventeenth week, the fine was removed with no explanation. Notice of the cancellation was posted on the board. If parents asked why the fines were removed, the principals were instructed to reply that the fine had been a trial for a limited time and that the results of this trial were now being evaluated.

A comparison with other fines in Israel may give an idea of the size of
the penalty that was introduced. A fine of NIS 10 is relatively small but not insignificant. In comparison, the fine for illegal parking is NIS 75; the fine for driving through a red light is NIS 1,000 plus penalties; the fine for not collecting the droppings of a dog is NIS 360. For many of these violations, however, detection and enforcement are low or, as in the case of dog dirt, nonexistent in practice. A baby-sitter earns between NIS 15 and NIS 20 per hour. The average gross salary per month in Israel at the time of the study was NIS 5,595.

</div>

</div>


<a href="javascript:showhide('wide')">The Data: Wide Format <span style="font-size:8pt;">(click to view)</span></a>

<div id="wide" style="display:none;">

The `late` Day Care Center data is shown here in the "wide data format". 

```{r, out.width=300}
# To get the "late" dataset into your R Console, 
# you need to click the green "play" arrow in 
# the top right corner of this gray R-Chunk. 
# Then type:
#   > View(late)
# in your R Console
late <- rbind(c(1, 37, 8, 8, 7, 6, 8, 9, 9, 12, 13, 13, 
                15, 13, 14, 16, 14, 15, 16, 13, 15, 17),
              c(2, 35, 6, 7, 3, 5, 2, 11, 14, 9, 16, 12, 
                10, 14, 14, 16, 12, 17, 14, 10, 14, 15),
              c(3, 35, 8, 9, 8, 9, 3, 5, 15, 18, 16, 14, 
                20, 18, 25, 22, 27, 19, 20, 23, 23, 22),
              c(4, 34, 10, 3, 14, 9, 6, 24, 8, 22, 22, 19, 
                25, 18, 23, 22, 24, 17, 15, 23, 25, 18),
              c(5, 33, 13, 12, 9, 13, 15, 10, 27, 28, 35, 
                10, 24, 32, 29, 29, 26, 31, 26, 35, 29, 28),
              c(6, 28, 5, 8, 7, 5, 5, 9, 12, 14, 19, 17, 
                14, 13, 10, 15, 14, 16, 6, 12, 17, 13),
              c(7, 35, 7, 10, 12, 6, 4, 13, 7, 8, 
                5, 12, 3, 5, 6, 13, 7, 4, 7, 10, 4, 6),
              c(8, 34, 12, 9, 14, 18, 10, 11, 6, 
                15, 14, 13, 7, 12, 9, 9, 17, 8, 5, 11, 8, 13),
              c(9, 34, 3, 4, 9, 3, 3, 5, 9, 5, 2, 
                7, 6, 6, 9, 4, 9, 2, 3, 8, 3, 5),
              c(10, 32, 15, 13, 13, 12, 10, 9, 15, 15, 15, 
                10, 17, 12, 13, 11, 14, 17, 12, 9, 15, 13))
late <- data.frame(c(rep("Fine",6),rep("Control",4)), late)
colnames(late) <- c("Treatment","Center","No.ofChidren",paste("Week",1:20, sep=""))
knitr::kable(late)
```

</div>


<a href="javascript:showhide('long')">The Data: Long Format <span style="font-size:8pt;">(click to view)</span></a>


<div id="long" style="display:none;">

The `Late` Day Care Center data is shown here in the "long data format".

```{r}
# This code reshapes the data into "long" format.
# To get the "Late" dataset into your R Console, 
# you need to click the green "play" arrow in 
# the top right corner of this gray R-Chunk. 
# Then type:
#   > View(Late)
# in your R Console
Late <- reshape(late,
                varying = paste("Week",1:20, sep=""), 
                v.names = "No.ofLateChildren",
                timevar = "Week", 
                times = 1:20, 
                idvar = "Center",
                new.row.names = 1:200,
                direction = "long")
knitr::kable(Late)
```

</div>



### Questions and Hypotheses 

The significance level will be set at 

$$
  \alpha = 0.05
$$
Since a Two-Way ANOVA test will be performed, there will not be three separate distributions of the significance level. 

<a href="javascript:showhide('question1')">Question 1<span style="font-size:8pt;">&nbsp;(click to view)</span></a>

<div id="question1" style="display:none;">

Does Treatment type (Control versus Fine Group) affect the number of children who were late?

$$
  H_0: \mu_C = \mu_F = \mu
$$
$$
  H_a: \mu_i \neq \mu \ \text{for at least one}\ i\in\{1=A,2=B\}
$$
The null hypothesis shows the mean of the control treatment (C), equals the mean of the fine treatment (F). 

The alternative hypothesis shows at least one of two sample means is different.

</div>


<a href="javascript:showhide('question2')">Question 2<span style="font-size:8pt;">&nbsp;(click to view)</span></a>

<div id="question2" style="display:none;">

Does Time Period (Before, During, or After the Fine) affect the number of children who were late?

$$
  H_0: \mu_B = \mu_D = \mu_A = \mu
$$
$$
  H_a: \mu_i \neq \mu \ \text{for at least one}\ i\in\{1=L,2=M,3=H\}
$$
The null hypothesis shows the mean of the Before period (B),  equals the mean of the During period (D), equals the mean of the After period (A). 

The alternative hypothesis shows at least one of three sample means is different.

</div>


<a href="javascript:showhide('question3')">Question 3<span style="font-size:8pt;">&nbsp;(click to view)</span></a>

<div id="question3" style="display:none;">

Does Time Period change for different treatments?
$$
  H_0: \text{The effect of periods is the same for both treatments.}
$$

$$
  H_a: \text{The effect of periods is not the same for both treatments.}
$$
The null hypothesis shows the effect of periods (Before, During, and After) is the same for both treatments (Control and Fine). 

The alternative hypothesis shows the effect of periods (Before, During, and After) is not the same for both treatments (Control and Fine).  


</div>


## Analysis

The Two-Way ANOVA is performed by using this model.  

$$
  Y_{ijk} = \mu + \alpha_i + \beta_j + \alpha\beta_{ij} + \epsilon_{ijk}
$$
$\mu$ is the grand mean

$\alpha_i$ is the Treatment factor with levels $C=1$ and $F=2$

$\beta_j$ is the Period factor with levels $B=1$, $D=2$, and $A=3$

$\alpha\beta_{ij}$ is the interaction of the two factors which has $2\times3=6$ levels

$\epsilon_{ijk} \sim N(0,\sigma^2)$ is the error term. 



```{r}
Late <- read.csv("../Data/Late.csv", header=TRUE)
Late$Period <- mapvalues(Late$Week, 1:20,c(rep("Before", 4), rep("During", 12), rep("After", 4)))
Late.aov <- aov(No.ofLateChildren ~ Period + Treatment + Period:Treatment, data=Late)
pander(summary(Late.aov))
par(mfrow=c(1,2))
plot(Late.aov, which=1:2, pch=16)
# Using the cut function! 
# cut(x, breaks=c(17, 34, 64, Inf), labels=c("18-34", "35-64", "65+"))
```

### Requirements for the Tests

On the left, the residual plot demonstrates the constant variances for each group are somewhat questionable because the spread of the vertical data sets is becoming larger as the fitted values increase.  On the right, the QQ plot shows the errors are normally distributed because the points follow an approximately straight line.  Even though the change in variance is questionable, the spread is not large enough to be revoked for the Two-Way ANOVA Test. The validity of the tests is sufficient. 

###Results of the Tests

<a href="javascript:showhide('question11')">Question 1<span style="font-size:8pt;">&nbsp;(click to view)</span></a>

<div id="question11" style="display:none;">


Comparing the Treatments, there is sufficient evidence to conclude at least one of the sample means is different from the others.   $(p=4.925 e -12 < \alpha)$. 

</div>

<a href="javascript:showhide('question22')">Question 2<span style="font-size:8pt;">&nbsp;(click to view)</span></a>

<div id="question22" style="display:none;">

Comparing the Periods, there is sufficient evidence to conclude at least one of the sample means is different from the others.  $(p=5.456 e -6 < \alpha)$. 

</div>

<a href="javascript:showhide('question33')">Question 3<span style="font-size:8pt;">&nbsp;(click to view)</span></a>

<div id="question33" style="display:none;">


Comparing the effect of Periods on Treatments, there is sufficient evidence to conclude the effect of Periods is not the same for both Treatments.  $(p=4.162 e -6 < \alpha)$. 

</div>


## Graphical Summary

```{r, message=FALSE, warning=FALSE}
Late$Period <- factor (Late$Period, levels = c("Before", "During", "After"))

xyplot( No.ofLateChildren ~ Period, data = Late, groups=Treatment, main="Number of Late Children During Time Periods", jitter.x=TRUE, jitter.y=TRUE, type=c("p", "a"), auto.key=list(corner=c(0,1))) 
```

## Interpretation 

<a href="javascript:showhide('question111')">Question 1<span style="font-size:8pt;">&nbsp;(click to view)</span></a>

<div id="question111" style="display:none;">


In regards to Treatment, the graph shows an equal distribution of Treatments in the Before Period, but in the During and After Periods, the majority of data points from the Fine Treatment is at the top, concluding there is a difference in the Treatment groups. 

</div>

<a href="javascript:showhide('question222')">Question 2<span style="font-size:8pt;">&nbsp;(click to view)</span></a>

<div id="question222" style="display:none;">

In regards to Period, the graph shows the average number of 10 children who were late in the Before Period.  Meanwhile, the number of late children increase to an average number of 18 children who were late in the During and After Periods.  Thus concluding there is a difference in the Periods.  

</div>

<a href="javascript:showhide('question333')">Question 3<span style="font-size:8pt;">&nbsp;(click to view)</span></a>

<div id="question333" style="display:none;">

In regards to Treatment and Period, the blue line indicates an approximately steady number of children who were late throughout all the periods of time versus the pink line which indicates a rapid increase in Before and During, and continues After the fine was revoked, showing there was an effect of Periods was not the same for both Treatments.  

</div>


**Conclusion** 

There is a relationship between Treatment and Period.  As the Period time increased there was a remarkable difference in the number of children who were late in the Fine group in comparison to the baseline, Control group.  

Interestingly enough, after the fine was withdrawn, the number of late children did not decrease.  Instead, the number actually slightly increased.  Did the fine change how these parents viewed their obligations?  Maybe the  parents felt it was okay to be late to pick up their children when they were paying a fine.  

A further study could be performed over a longer period of time to focus on the behavior of people taking advantage of a priviledge, then have it withdrawn for a time, then to receive the priviledge again to measure their behaviors to see if the results are consistent with the results from this study. 
