---
title: "Thinking Aloud"
output: 
  html_document:
    theme: flatly
    code_folding: hide
    toc: true 
    toc_float: true
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r, include=FALSE}
library(mosaic)
library(car)
library(DT)
library(pander)
library(yarrr)

Baumann <- read.csv("../Data/READING.csv", header=TRUE)


Baumann$Difference2 <- Baumann$post.test.2 - Baumann$pretest.2
```

```{r, eval=FALSE}
#Play the above chunk and this one in your Console to access the data
View(Baumann)
?Baumann
```

## Background

Thinking aloud is known to be beneficial, but is there a method that works better than traditional instruction?  A study was performed to 4th graders in the midwestern United States testing three different instructional methods to see if students improved after each method was implemented.  Details of the study are shown below. 


"The [empirical] background for the experiment draws from research on comprehension [monitoring] and the [think aloud] procedure."

"This study investigated the effectiveness of explicit instruction in 'think aloud' as a means to promote elementary students' comprehension monitoring abilities" [(Baumann & Jones, 1992)](http://jlr.sagepub.com/content/24/2/143.full.pdf).


<a href="javascript:showhide('abstract')">Details of the Study<span style="font-size:8pt;">&nbsp;(click to view)</span></a>

<div id="abstract" style="display:none;">

The following explanation of the experiment comes directly from the original research article [(Baumann & Jones, 1992)](http://jlr.sagepub.com/content/24/2/143.full.pdf).

<div style="padding-left:30px; padding-right:30px;">

Fourth-grade students from an elementary school in a rural midwestern community participated in the study. The school had adopted a team teaching plan, and three fourth-grade teachers shared an open classroom area. The teachers grouped the students for reading instruction across the three classes according to reading ability. One teacher each worked with a high [ability reading group], [one teacher worked with a] middle [ability group], and [the other teacher worked with the] low reading [ability] group. All groups used basal reading materials and were formed at the beginning of the school year on the basis of standardized achievement test data, prior performance in the basal reading program, and teacher judgment.

Sixty-eight of the 72 students in the fourth-grade team participated in the study; the four nonparticipating students were mainstreamed special education students for whom the instructional materials were considered to be too frustrating to read. The 68 students were assigned randomly to one of the three experimental groups (Strat = 23 students; DRTA = 23 students; Basal = 22 students). Because of illness and transfer to another school, incomplete data were obtained for one subject each from the Strat and DRTA groups. Thus, the final sample consisted of 66 students, 32 girls and 34 boys.

Posttests 1-3 were group-administered, quantitative measures obtained
for all students in the sample. Posttest 1 was an error detection test. Posttest 2 was a comprehension monitoring questionnaire. Posttest 3 was a modified cloze test. To statistically account for students'
preexperimental differences in comprehension monitoring abilities, two pretests were constructed, administered, and used as covariates in data analyses for Posttests 1-3.

Pretest 1. The first pretest involved an error detection task designed to evaluate students' preintervention ability to monitor their comprehension.... Students were provided the following directions which were printed on the pretest and read aloud by the experimenter: 'Read the story carefully. Try to understand it as you read. There will be 16 sentences that do not belong in the story. Look for them as you read. Ask yourself, "What does not make sense in this story?" Underline the 16 sentences that do not make sense. Check your work when you are finished. If you cannot read a
word, raise your hand, and we will help you.' Preceding Pretest 1, students were given a brief practice exercise that required them to identify obvious intruded sentences in two short texts. Pretest 1 was scored by calculating the number of correctly identified intruded sentences.

Pretest 2. The second pretest queried students about the strategies they believed to be useful in promoting their understanding of stories. In this pretest, modeled after an instrument developed by Schmitt (1988, 1990), the students were presented with 15 multiple-choice items like the following:

<div style="padding-left:15px;">
When I read, it is a good idea to:

A. sound out words I don't know.

B. make some guesses about what will happen in the story.

C. make a list of all the details in the story.

D. look up new words in the dictionary.

</div>

Posttest 1: Error detection test. Posttest 1 was the same in form as Pretest 1: with 16 intruded sentences lexically consistent at a local level but semantically inconsistent at a global level. 

Posttest 2: Comprehension monitoring questionnaire. Posttest 2, consisting of 18 items, was a slightly expanded version of Pretest 2. 

Posttest 3: Degrees of Reading Power. The Degrees of Reading Power (DRP)
(1986) is a standardized, commercially produced instrument that employs a modified cloze format (words are deleted from a selection and for each omission the student selects the most appropriate word from a list of five). According to the authors, DRP tests "are holistic measures of how well the messages within text are understood. . . . DRP tests focus measurement on determining how well students process or construct meaning from paragraphs as they read through a selection" (DRP Handbook, 1986, p. 1). Thus, the DRP, at one level, is a general reading comprehension assessment tool.

</div>

The data for this study are contained in the `Baumann` dataset.

</div>


### Question and Hypotheses

The significance level will be set at 

$$
  \alpha = 0.05
$$

Since it is more affective to measure the pretest against the post test, a Kruskal-Wallis Test will be performed to analyze the differences between the post and pre test of the second test.  


<a href="javascript:showhide('question')">Question<span style="font-size:8pt;">&nbsp;(click to view)</span></a>

<div id="question" style="display:none;">

Question: For the pre/post test 2, is there a difference in the instructional methods?

$$
  H_0: \text{All the instructional methods have the same distribution.}
$$ 

The null hypothesis is interpreted as all the instructional methods: Basal, DRTA, and Strat are from the same distribution. 

$$
  H_a: \text{At least one of the instructional methods is stochastically different.}
$$

The alternative hypothesis is interpreted as at least one of the instructional methods: Basal, DRTA, or Strat is stochastically different. (generally accepted to be true) 


</div>

## Analysis 

All three of the instructional methods were compared. 

* Basal: Standard teaching method
* DRTA:  Directed Reading Thinking Aloud 
* Strat: Another innovative method

A Kruskal-Wallis test was conducted.  See below for the results. 


```{r}
pander(kruskal.test(Difference2 ~ group, data=Baumann))
```

This pirate plot below shows a comparison of the hypotheses to the instructional methods:  Basal, DRTA, and Strat.

```{r}
pirateplot(formula = Difference2 ~ group,
          data = Baumann,
          theme = 3,
          pal = "appletv",
          bean.f.o = .5,
          back.col = grey(.97),
          quant = c(.1, .9), # 10th and 90th quantiles
          quant.col = "black", # Black quantile lines
          main = "Differences in Post and Pre Test 2", 
          xlab = "Instructional method", 
          ylab = "Score")
```

This 5 number summary below represents the distribution of the three instructional methods. 

```{r}
pander(favstats(Difference2 ~ group, data=Baumann))
```

### Results of the Test

Comparing instructional methods, there is sufficient evidence to conclude at least one of the instructional methods is different from the others.

```{r, echo=FALSE}
##boxplot(Difference2 ~ group, data=Baumann)
# Boxplot tests to see which Null and Alternative hypothesis to use.  Since each group does not have the same shape/distributions, we will use the Null and alternative of stochastical dominance.
```


## Interpretation

The pirate plot demonstrates one of the instructional methods is different (infering the Strat method) and is confirmed by the five number summary showing Basal and DRTA's median as 0 and Strat's median as 3.  This confirms the alternative hypothesis that at least one of the instructional methods is different from the others.  

### Conclusion

In addition to the interpretation, there are also a few concepts to consider.  For each instructional method, there were only 22 participants for each, which is a small sample size.  It would be valuable to perform another test to check consistency by increasing the sample size of each instructional method.  

Also important to note, the post test had three more questions added to it then the pre test.  Ideally it would be best to keep the amount of questions the same for consistency.  

There are possibly multiple ways to enhance thinking aloud in students.  Further testing would be beneficial in confirming the validity of the tests. 
